\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{microtype}

\title{Beyond Scalar CISPO: Causal-Token Credit Allocation for Reasoning RL}
\author{}
\date{}

\begin{document}
\maketitle

\begin{abstract}
Long-chain reasoning reinforcement learning depends on stable policy-gradient
updates over token sequences, but current clipped objectives usually broadcast
the same sequence advantage and the same trust-region geometry to every
generated token. This proposal studies a family of CISPO extensions that keep
MiniMax's clipped importance-sampling policy optimization (CISPO) as the
stability backbone, while addressing what CISPO does not solve: token-level
credit allocation under sparse sequence rewards. The first AD-CISPO prototype
redistributed the upper clipping radius according to token saliency, but
preliminary runs suggest that bound-only adaptation is a weak lever when few
policy ratios reach the adaptive bound. The revised next-generation proposal is
Credit-Weighted CISPO (CW-CISPO): a detached, mean-preserving token-weighted
CISPO objective whose weights are computed from target-policy causal-tangent
saliency, optionally gated by token entropy. Adaptive bounds remain as a
guardrail and ablation rather than the sole use of saliency.
\end{abstract}

\section{Problem and Motivation}

Reinforcement learning has become a practical way to elicit mathematical and
long-form reasoning from language models, especially when correctness can be
checked by rule-based or programmatic rewards. PPO-style RLHF established
clipped policy-gradient training as a scalable alignment method
\cite{schulman2017ppo,ouyang2022instructgpt}, while GRPO removed the critic by
using group-relative advantages and has been widely used for mathematical
reasoning \cite{shao2024deepseekmath,deepseek2025r1}. More recent systems such
as DAPO and MiniMax-M1 focus on stabilizing long-chain reasoning RL at scale
\cite{yu2025dapo,minimax2025m1}.

The unresolved problem is broader than scalar clipping: current outcome-reward
RL largely broadcasts a sequence-level advantage to every generated token. In
chain-of-thought-style reasoning, this is unlikely to be true. A token that
introduces an intermediate quantity, selects an operation, or anchors a later
conclusion may matter more than a formatting token or locally redundant phrase.
CISPO fixes one important failure of PPO-style clipping by preserving dense
gradients under clipped IS weights, but it still leaves token credit largely
uniform. A single global clipping radius and a uniform token loss weight
therefore create an undesirable tradeoff:
\begin{itemize}[leftmargin=*]
  \item if the radius is tight, important reasoning tokens may be prevented
  from moving enough when the reward signal identifies a better trajectory;
  \item if the radius is loose, low-value or unstable tokens receive the same
  update corridor, increasing variance and drift;
  \item if the same sequence advantage is broadcast to all tokens, deterministic
  filler tokens can absorb gradient mass that should be spent on uncertain or
  causally important reasoning decisions;
  \item if PPO/GRPO surrogate clipping suppresses token updates, the model can
  lose useful gradient signal exactly when the sampled policy has moved outside
  the trust corridor.
\end{itemize}

CISPO directly targets the third issue by clipping the importance-sampling
(IS) weight itself and using the detached clipped weight to scale
$\log \pi_\theta$, so clipped tokens still contribute gradients
\cite{minimax2025m1}. This proposal asks whether the remaining scalar-budget issue
and the more general token-credit issue can be addressed without abandoning
CISPO's dense-gradient design.

\section{Existing Solutions and Limitations}

\paragraph{PPO and GRPO.}
PPO clips the policy-ratio surrogate to control policy updates while allowing
multiple optimization epochs per batch \cite{schulman2017ppo}. GRPO adapts
this family of methods to reasoning models by computing advantages from a
group of sampled responses rather than a learned critic
\cite{shao2024deepseekmath}. These methods are simple and effective, but their
token-level clipping rules still use fixed trust-region hyperparameters.

\paragraph{DAPO and asymmetric clipping.}
DAPO improves long-chain reasoning RL with techniques such as a higher upper
clip, dynamic sampling, token-level policy-gradient loss, and overlong reward
shaping \cite{yu2025dapo}. This is an important baseline because it shows that
the upper clipping corridor matters for reasoning RL. However, DAPO still
treats the upper corridor as a sequence-wide or globally configured scalar,
not as a budget to be distributed across tokens.

\paragraph{CISPO.}
MiniMax-M1 proposes CISPO, which clips IS weights rather than clipping away
token updates and reports improved RL efficiency in large-scale reasoning
training \cite{minimax2025m1}. CISPO is the strongest starting point for this
proposal because it preserves per-token gradient density. Its limitation for
this project is that it still broadcasts the same sequence advantage across
tokens and uses a scalar IS clipping geometry.

\paragraph{Token-level credit assignment and reward redistribution.}
The broader RL literature has long treated delayed rewards as a credit
assignment problem. RUDDER, for example, redistributes delayed returns through
contribution analysis so that future rewards become easier to predict
\cite{arjona2019rudder}. In LLM alignment, process supervision and process
reward models explicitly replace pure outcome supervision with step-level
feedback \cite{lightman2023verify}. Recent token-level preference methods,
including TDPO, TIS-DPO, SePO, T-REG, and TI-DPO, also argue that sequence-level
preference optimization hides important token-level structure
\cite{zeng2024tdpo,liu2025tisdpo,yang2025sepo,zhou2025treg,yang2025tidpo}.
These works are important prior art for the revised proposal: they show that
token-level weighting and token selection are active directions, so the novelty
cannot simply be ``use token weights.'' The gap is instead online
verifiable-reward RL built specifically on CISPO's clipped-IS gradient, without
external token reward labels, pairwise preference data, or a separately trained
oracle selector.

\paragraph{Entropy and capacity views of RLVR credit.}
Recent RLVR analysis frames token credit as a reward-conditioned shift from the
behavior policy toward a hindsight posterior, with token entropy upper-bounding
where hindsight credit can reside \cite{he2026hindsightcredit}. This suggests
that saliency alone is incomplete: a token can be locally objective-sensitive
yet carry little learnable credit if it is already deterministic. A stronger
CISPO extension should therefore test saliency-only weights against
entropy-gated weights that focus update mass where the policy has both
objective sensitivity and uncertainty capacity.

\paragraph{Attention-based saliency.}
Transformers route information through attention mechanisms
\cite{vaswani2017attention}, and attention patterns can provide useful
structural signals. However, attention weights should not be treated as direct
causal explanations of model behavior \cite{jain2019attention}. Attention
sinks are a concrete example: initial or special tokens may receive high
attention for non-semantic reasons \cite{xiao2024streamingllm}. AD-CISPO
therefore uses frozen-reference attention only as an update-budget heuristic,
not as a claim about human-interpretable explanation, and explicitly masks
prompt and sink tokens.

\paragraph{Mechanistic attribution and gradient saliency.}
Mechanistic interpretability distinguishes observational signals such as
attention patterns from intervention-based methods such as activation patching
and causal tracing \cite{meng2022rome,wang2022ioi}. Attribution patching
approximates activation-patching effects with first-order gradients, making it
more scalable but also susceptible to local-linearity failures
\cite{syed2023attribution,kramar2024atpstar}. Classical attribution methods
such as Integrated Gradients, DeepLIFT, and SmoothGrad further show that raw
gradient saliency is scale-sensitive and noisy unless it is anchored to a
baseline, averaged, or otherwise regularized
\cite{sundararajan2017ig,shrikumar2017deeplift,smilkov2017smoothgrad}. These
lessons motivate bounded, detached, mean-preserving weights and local smoothing:
causal-tangent saliency should be treated as an optimization signal to ablate,
not as a faithful explanation by default.

\section{Research Gap}

Existing reasoning-RL objectives answer two separate questions only partially:
\begin{enumerate}[leftmargin=*]
  \item How can policy-gradient updates remain stable under off-policy drift?
  PPO, GRPO, DAPO, and CISPO address this through clipped ratios or clipped
  IS weights.
  \item Which generated tokens deserve a larger or smaller learning signal?
  Existing online RLVR methods largely still broadcast sequence-level rewards,
  even when they improve clipping, sampling, or normalization.
\end{enumerate}

The gap is a CISPO-native token-credit objective for language-model RL: one
that keeps CISPO's detached clipped-IS gradient as the stability mechanism, but
adds a bounded, falsifiable token-credit allocation layer derived from the
current policy's own computation on the sampled trajectory. This is different
from process reward modeling, token-level DPO, or token selection in
preference data because the signal is computed online from the RL rollout and
used inside the CISPO gradient itself.

\section{Proposed Solution}

The revised proposal treats CISPO as a stable base objective and studies three
increasingly strong ways to add token credit:
\begin{enumerate}[leftmargin=*]
  \item \textbf{AD-CISPO-Bound:} the original conservative variant, where
  saliency only redistributes the upper IS clipping radius.
  \item \textbf{Credit-Weighted CISPO (CW-CISPO):} the lead next-generation
  variant, where saliency becomes a detached, bounded, mean-one multiplier on
  each token's CISPO gradient while the scalar CISPO clip remains a guardrail.
  \item \textbf{Hybrid CW-AD-CISPO:} an ablation where saliency controls both
  the token loss weight and the upper IS clipping radius, allowing the project
  to test whether credit allocation or trust-region allocation is the useful
  mechanism.
\end{enumerate}

The lower IS bound remains inactive by default to match the MiniMax-style CISPO
baseline. Each generated token receives a multiplier computed from a saliency
extractor, and the multipliers are normalized so their mean over valid
generated tokens is $1$. This preserves the average per-sequence gradient scale
or clipping budget while changing where the budget is spent. The first
implementation used frozen-reference future-attention in-degree; the
next-generation implementation should use target-policy causal-tangent saliency
because attention in-degree can be dominated by sink tokens and because the
target policy is the computation graph actually being optimized.

\paragraph{Core hypothesis.}
For reasoning tasks with sparse sequence-level rewards, CISPO's dense-gradient
property is necessary but not sufficient: learning efficiency is limited by
uniform token-credit broadcast. Bounded, mean-preserving causal-token weights
should improve reward efficiency relative to scalar CISPO if they move gradient
mass toward objective-sensitive and uncertainty-bearing reasoning tokens. If
shuffled weights, entropy-only weights, or uniform multipliers match the gain,
then the method has not demonstrated a saliency-specific advantage.

\paragraph{What is novel.}
The proposal's novelty is not the use of one particular saliency heuristic by
itself, nor merely the fact of token weighting. The novel contribution is the
combination of:
\begin{enumerate}[leftmargin=*]
  \item CISPO-compatible token-credit allocation that preserves the detached
  clipped-IS dense-gradient structure rather than replacing it with a
  preference-objective, process reward model, or PPO surrogate;
  \item mean-preserving weight and multiplier normalization, so CW-CISPO
  reallocates gradient mass rather than increasing the total average
  per-sequence learning rate;
  \item adaptive upper-clip redistribution as a controlled ablation and
  guardrail, not the only way saliency affects optimization;
  \item an explicit saliency interface, where attention in-degree is a cheap
  baseline, causal-tangent saliency is the first objective-sensitive
  replacement, and entropy-gated causal saliency is the first capacity-aware
  variant;
  \item sink-token diagnostics and guards, so punctuation or boilerplate
  attention sinks cannot silently absorb the adaptive learning budget;
  \item falsifiable ablations that separate saliency from extra variance,
  entropy gating, asymmetric clipping, and reference-model overhead.
\end{enumerate}

\paragraph{Novelty boundary after literature review.}
The proposal should not claim to invent token-level credit assignment, token
selection, gradient attribution, or reward redistribution. Those ideas already
exist in RUDDER, process supervision, token-level DPO variants, selective
preference optimization, and mechanistic attribution. The narrower novelty
claim is: \emph{online, verifier-reward CISPO with detached clipped-IS weights,
augmented by bounded mean-one token-credit weights derived from the target
policy's causal-tangent sensitivity and optionally gated by entropy}. If a
prior method is found that already combines all of these elements, the fallback
novelty becomes empirical: a controlled comparison showing whether
CISPO-specific credit weighting improves reasoning RL where scalar CISPO is
already strong.

\section{Notation}

\begin{center}
\small
\begin{tabular}{p{0.20\linewidth}p{0.48\linewidth}p{0.22\linewidth}}
\toprule
Symbol & Meaning & Shape or domain \\
\midrule
$q$ & prompt/question & text \\
$i$ & index of a sampled response within a prompt group & $i \in \{1,\ldots,G\}$ \\
$o_i$ & response $i$ in a sampled group & token sequence \\
$t$ & token position in response $i$ whose saliency, IS weight, or multiplier is being computed & $t \in M_i$ \\
$j$ & later token position in response $i$ that attends back to position $t$ & $j \in M_i$, $j > t$ \\
$u$ & summation index over valid generated positions in response $i$ & $u \in M_i$ \\
$o_{i,t}$ & token at position $t$ of response $i$ & token id \\
$G$ & number of responses sampled per prompt & positive integer \\
$\pi_\theta$ & current policy being optimized & language model \\
$\pi_{\theta_{\mathrm{old}}}$ & rollout policy that produced the batch & language model \\
$\pi_{\mathrm{ref}}$ & frozen reference model for optional KL and saliency baselines & language model \\
$r_{i,t}(\theta)$ & IS weight
  $\frac{\pi_\theta(o_{i,t}|q,o_{i,<t})}
  {\pi_{\theta_{\mathrm{old}}}(o_{i,t}|q,o_{i,<t})}$ & positive scalar \\
$\hat{A}_i$ & group-relative sequence advantage for response $i$ & scalar \\
$M_i$ & valid generated-token action positions for response $i$ & token-position set \\
$S_{i,t}$ & raw saliency for token $o_{i,t}$ & nonnegative scalar \\
$H_{i,t}$ & token entropy under the rollout or target policy & nonnegative scalar \\
$C_{i,t}$ & optional capacity gate derived from entropy or uncertainty & nonnegative scalar \\
$m_{i,t}$ & normalized token multiplier used for adaptive clipping & nonnegative scalar, mean $1$ over $M_i$ \\
$w_{i,t}$ & detached normalized token-credit weight used in CW-CISPO & nonnegative scalar, mean $1$ over $M_i$ \\
$\epsilon^{IS}_{\mathrm{low}}$ & lower IS clipping radius & nonnegative scalar \\
$\epsilon^{IS}_{\mathrm{high}}$ & scalar upper IS clipping radius & positive scalar \\
$\epsilon^{IS}_{i,t,\mathrm{high}}$ & token-adaptive upper IS clipping radius & positive scalar \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Index conventions.}
For each prompt, the model samples $G$ responses; $i$ selects one of them.
Within response $i$, positions $t$, $j$, and $u$ all refer to token indices in
the same sequence, but play different roles: $t$ is the position whose saliency
or clipping budget is being assigned; $j$ is a \emph{later} position whose
attention query looks back at $t$; and $u$ is a dummy index used only when
averaging or summing over all valid generated positions in $M_i$.

\section{Baseline: MiniMax-Style CISPO}

The CISPO baseline starts from the off-policy REINFORCE objective with the IS
weight treated as a stop-gradient coefficient:
\begin{equation}
  J_{\mathrm{ISPG}}(\theta)
  =
  \mathbb{E}
  \left[
    \frac{1}{\sum_{i=1}^G |o_i|}
    \sum_{i=1}^G \sum_{t=1}^{|o_i|}
    \operatorname{sg}\!\left(r_{i,t}(\theta)\right)
    \hat{A}_i
    \log \pi_\theta(o_{i,t}\mid q,o_{i,<t})
  \right].
\end{equation}

CISPO clips the IS weight itself:
\begin{equation}
  \hat{r}_{i,t}(\theta)
  =
  \operatorname{clip}
  \left(
    r_{i,t}(\theta),
    1-\epsilon^{IS}_{\mathrm{low}},
    1+\epsilon^{IS}_{\mathrm{high}}
  \right),
\end{equation}
and optimizes
\begin{equation}
  J_{\mathrm{CISPO}}(\theta)
  =
  \mathbb{E}
  \left[
    \frac{1}{\sum_{i=1}^G |o_i|}
    \sum_{i=1}^G \sum_{t=1}^{|o_i|}
    \operatorname{sg}\!\left(\hat{r}_{i,t}(\theta)\right)
    \hat{A}_i
    \log \pi_\theta(o_{i,t}\mid q,o_{i,<t})
  \right].
\end{equation}

A faithful baseline for this proposal sets the KL coefficient to zero and
avoids an active lower IS-weight bound, following the MiniMax-M1 description of
CISPO \cite{minimax2025m1}. This isolates the effect of the upper clipping
budget.

\section{Bound-Only AD-CISPO Objective}

The original AD-CISPO variant replaces the scalar upper clipping radius with a
token-adaptive upper radius:
\begin{equation}
  \epsilon^{IS}_{i,t,\mathrm{high}}
  =
  \epsilon^{IS}_{\mathrm{high}} m_{i,t}.
\end{equation}
The clipped weight becomes
\begin{equation}
  \hat{r}^{\mathrm{AD}}_{i,t}(\theta)
  =
  \operatorname{clip}
  \left(
    r_{i,t}(\theta),
    1-\epsilon^{IS}_{\mathrm{low}},
    1+\epsilon^{IS}_{i,t,\mathrm{high}}
  \right),
\end{equation}
and the objective is
\begin{equation}
  J_{\mathrm{AD\mbox{-}CISPO}}(\theta)
  =
  \mathbb{E}
  \left[
    \frac{1}{\sum_{i=1}^G |o_i|}
    \sum_{i=1}^G \sum_{t=1}^{|o_i|}
    \operatorname{sg}\!\left(\hat{r}^{\mathrm{AD}}_{i,t}(\theta)\right)
    \hat{A}_i
    \log \pi_\theta(o_{i,t}\mid q,o_{i,<t})
  \right].
\end{equation}

The adaptive multiplier is computed from token saliency. Let $M_i$ denote the
set of generated-token action positions in response $i$ that are eligible for
adaptive clipping. It excludes prompt positions, padding, masked-out actions,
special tokens, and any explicit sink-token categories removed by the saliency
guard. Here $u$ ranges over all positions in $M_i$ to form the mean saliency in
the denominator, while $t$ is the position whose multiplier is being computed:
\begin{equation}
  \tilde{m}_{i,t}
  =
  \frac{S_{i,t}}
  {\frac{1}{|M_i|}\sum_{u \in M_i} S_{i,u}}.
\end{equation}
If all valid saliency values are zero or non-finite, $\tilde{m}_{i,t}$ falls
back to $1$ for valid tokens. If lower or upper multiplier bounds are used,
they must be followed by renormalization or projection so that
\begin{equation}
  \frac{1}{|M_i|}\sum_{t \in M_i} m_{i,t} = 1.
\end{equation}
A plain clamp is insufficient because it changes the total clipping budget.

\section{Credit-Weighted CISPO Objective}

The proposed next-generation objective uses saliency as a token-credit weight
rather than only as an upper-bound modifier. Starting from the CISPO clipped IS
weight $\hat{r}_{i,t}(\theta)$, define a detached nonnegative token-credit
weight $w_{i,t}$ satisfying
\begin{equation}
  \frac{1}{|M_i|}\sum_{t \in M_i} w_{i,t} = 1,
  \qquad
  w_{\min} \leq w_{i,t} \leq w_{\max}.
\end{equation}
CW-CISPO optimizes
\begin{equation}
  J_{\mathrm{CW\mbox{-}CISPO}}(\theta)
  =
  \mathbb{E}
  \left[
    \frac{1}{\sum_{i=1}^G |M_i|}
    \sum_{i=1}^G \sum_{t \in M_i}
    \operatorname{sg}\!\left(w_{i,t}\right)
    \operatorname{sg}\!\left(\hat{r}_{i,t}(\theta)\right)
    \hat{A}_i
    \log \pi_\theta(o_{i,t}\mid q,o_{i,<t})
  \right].
\end{equation}

This objective leaves CISPO's clipped-IS guardrail intact, but changes the
allocation of gradient mass across tokens. Mean-one normalization is critical:
without it, a run could improve merely because it increased the effective
learning rate. Detaching $w_{i,t}$ is also critical: the saliency pass should
decide how to allocate credit for the sampled trajectory, not create a second
implicit objective that backpropagates through the saliency extractor during
the policy update.

The simplest saliency-only weight is
\begin{equation}
  \tilde{w}_{i,t}
  =
  \left(
    \frac{S_{i,t}}
    {\frac{1}{|M_i|}\sum_{u \in M_i} S_{i,u} + \epsilon}
  \right)^\gamma,
\end{equation}
followed by bounded projection to mean $1$. The exponent $\gamma \in [0,1]$
controls how sharply the weight follows saliency. A conservative first run
should use $\gamma=1$ with $w_{\min}=0.25$ and $w_{\max}=2.0$, matching the
existing AD-CISPO multiplier range.

\paragraph{Entropy-gated causal-token credit.}
Recent RLVR credit analysis suggests that entropy marks where hindsight credit
can reside, while saliency marks where the current objective is locally
sensitive \cite{he2026hindsightcredit}. A capacity-aware variant should
therefore combine both:
\begin{equation}
  \tilde{w}^{\mathrm{cap}}_{i,t}
  =
  \left(S_{i,t}+\epsilon\right)^\gamma
  \left(C_{i,t}+\epsilon\right)^\eta,
  \qquad
  C_{i,t}
  =
  \frac{H_{i,t}}
  {\frac{1}{|M_i|}\sum_{u \in M_i} H_{i,u}+\epsilon}.
\end{equation}
Here $H_{i,t}$ can be the rollout-policy or target-policy entropy for the next
token distribution at position $t$. The ablation set should include saliency
only ($\eta=0$), entropy only ($\gamma=0$), and the product gate
($\gamma,\eta>0$). This prevents the proposal from confusing objective
sensitivity with learnable credit capacity.

\paragraph{Hybrid clipping and weighting.}
The hybrid variant uses one multiplier for credit weights and another for the
upper clipping radius:
\begin{equation}
  J_{\mathrm{Hybrid}}(\theta)
  =
  J_{\mathrm{CW\mbox{-}CISPO}}(\theta)
  \quad \text{with} \quad
  \epsilon^{IS}_{i,t,\mathrm{high}}
  =
  \epsilon^{IS}_{\mathrm{high}}m_{i,t}.
\end{equation}
This is intentionally not the first hypothesis. Preliminary AD-CISPO runs show
that upper-bound adaptation affects only the subset of tokens whose ratios
approach the bound. CW-CISPO is a stronger use of saliency because every token
gradient is reallocated while CISPO's scalar clip still limits off-policy
variance.

\section{Saliency Estimation}

The first AD-CISPO prototype used exact frozen-reference future-attention
in-degree. It remains a useful baseline because it is cheap relative to
target-policy causal-tangent saliency, but it should no longer be treated as
the primary extractor. For a selected reference layer $\ell$ and head $h$, let
$a^{(\ell,h)}_{i,j,t}$ denote the causal attention probability from the query at
token position $j$ to the key/value position $t$. Position $t$ is the saliency
target; position $j$ is each later valid generated token that attends to $t$.
The saliency assigned to token $o_{i,t}$ is
\begin{equation}
  S_{i,t}
  =
  \frac{1}{|L_{\mathrm{top}}|}
  \sum_{\ell \in L_{\mathrm{top}}}
  \frac{1}{H_\ell}
  \sum_{h=1}^{H_\ell}
  \sum_{\substack{j \in M_i \\ j > t}}
  a^{(\ell,h)}_{i,j,t}.
\end{equation}
Thus a token is salient when later valid generated tokens attend to it under
the frozen reference model. The strict condition $j>t$ prevents self-attention
from inflating saliency. In this estimator, $M_i$ is used both as the source
query set and target key set, so prompt, padding, special-token, and explicitly
guarded sink-token positions do not participate in the saliency budget.

The implementation computes this quantity blockwise from reference-model
post-RoPE query and key states. This avoids materializing a full batch of
$T \times T$ attention matrices while preserving exact softmax attention
weights for the selected layers, including causal, padding, and sliding-window
masks.

A cheaper ablation is frozen-reference key-vector norm:
\begin{equation}
  S^{\mathrm{KV}}_{i,t}
  =
  \frac{1}{|L_{\mathrm{top}}|}
  \sum_{\ell \in L_{\mathrm{top}}}
  \left\|k^{(\ell)}_{i,t}\right\|_2.
\end{equation}
It should be reported as a proxy baseline, not as the primary saliency
definition.

\section{Mathematical Failure Mode: Attention Sinks versus Causal Importance}

The future-attention in-degree estimator is best understood as an
attention-centrality score:
\begin{equation}
  S_{i,t}
  =
  \sum_{\ell,h}
  \sum_{\substack{j \in M_i \\ j > t}}
  a^{(\ell,h)}_{i,j,t}.
\end{equation}
It answers the question ``how much do later generated tokens look back at
position $t$?'' A CISPO credit-allocation method, however, ultimately needs a
different quantity:
how much changing the representation at position $t$ would change the
policy-gradient objective. These two quantities can diverge because attention
mass is not a causal explanation of model behavior \cite{jain2019attention}.

In a transformer attention layer, the value mixed into query position $j$ is
\begin{equation}
  z_{i,j}
  =
  \sum_{t \leq j} a_{i,j,t} V_{i,t}.
\end{equation}
A large coefficient $a_{i,j,t}$ says that position $j$ reads from position
$t$, but it does not imply that $V_{i,t}$ changes the final answer, reward, or
policy-gradient update. This distinction matters for attention sinks
\cite{xiao2024streamingllm}: early separator or formatting tokens can attract
many small attention weights from later positions for non-semantic reasons.

For example, suppose a response begins with a common reasoning preamble:
\begin{quote}
\texttt{Okay, let's try to figure out ...}
\end{quote}
Let $t_c$ be the comma position. If many future layer-head-query terms assign a
small amount of attention to this comma, the accumulated in-degree can dominate:
\begin{equation}
  a^{(\ell,h)}_{i,j,t_c} \approx 0.02,
  \qquad
  400 \text{ contributing terms}
  \quad\Rightarrow\quad
  S_{i,t_c} \approx 8.0.
\end{equation}
A genuinely reasoning-critical token may receive stronger attention from fewer
positions:
\begin{equation}
  20 \text{ contributing terms} \times 0.10
  \quad\Rightarrow\quad
  S_{i,t_{\mathrm{math}}} \approx 2.0.
\end{equation}
Under attention centrality, the comma receives the larger score even though it
may not be useful for solving the problem.

This creates a direct optimizer-level failure. AD-CISPO converts saliency into
an upper clipping budget:
\begin{equation}
  \epsilon^{IS}_{i,t,\mathrm{high}}
  =
  \epsilon^{IS}_{\mathrm{high}} m_{i,t}.
\end{equation}
If the comma sink receives $m_{i,t_c}=2$ and
$\epsilon^{IS}_{\mathrm{high}}=0.2$, its upper IS bound becomes
\begin{equation}
  1+\epsilon^{IS}_{i,t_c,\mathrm{high}}
  =
  1+0.2 \cdot 2
  =
  1.4.
\end{equation}
Thus the largest adaptive trust region can be allocated to punctuation in a
boilerplate prefix rather than to the tokens whose representations influence
the rewarded reasoning trajectory.

\paragraph{Causal-tangent saliency.}
A better first-generation replacement is an attribution-patching-style
first-order estimator, which approximates causal intervention effects with a
single backward pass \cite{meng2022rome,wang2022ioi,syed2023attribution}.
Define a teacher-forced scalar objective over sampled completions:
\begin{equation}
  F(\theta)
  =
  \sum_i
  |\hat{A}_i|
  \frac{
    \sum_{t \in M_i}
    \log \pi_\theta(o_{i,t}\mid q_i,o_{i,<t})
  }{|M_i|}.
\end{equation}
Let $h^{(\ell)}_{i,t}$ be the residual-stream activation at selected layer
$\ell$ and generated position $t$. A first-order Taylor expansion gives
\begin{equation}
  F(h^{(\ell)}_{i,t}+\Delta)
  -
  F(h^{(\ell)}_{i,t})
  \approx
  \nabla_{h^{(\ell)}_{i,t}} F \cdot \Delta.
\end{equation}

The word ``causal'' is used in the intervention sense, not as a claim that the
score is a complete causal proof. The ideal causal question is: if one could
intervene on the hidden state at token position $t$ while holding the rest of
the computation fixed as much as possible, how much would the scalar objective
$F$ change? Exact activation patching answers this by replacing or perturbing
$h^{(\ell)}_{i,t}$ and rerunning the downstream computation, but doing this for
every token, layer, and rollout is too expensive inside RL training. The
first-order term above is the tangent-space approximation to that intervention:
it estimates the local effect of an infinitesimal do-intervention
$\operatorname{do}(h^{(\ell)}_{i,t}\leftarrow h^{(\ell)}_{i,t}+\Delta)$ on
$F$. Thus the method is ``causal'' because it estimates the effect of changing
an internal state on the objective, and ``tangent'' because it uses the local
linearization rather than an explicit finite intervention.

This suggests a token saliency score
\begin{equation}
  S^{\mathrm{CT}}_{i,t}
  =
  \sum_{\ell \in L_{\mathrm{top}}}
  \frac{
    \left|
      \nabla_{h^{(\ell)}_{i,t}} F
      \cdot
      \left(
        h^{(\ell)}_{i,t}
        -
        \frac{1}{|M_i|}
        \sum_{u \in M_i} h^{(\ell)}_{i,u}
      \right)
    \right|
  }{\sqrt{d_{\mathrm{model}}}}.
\end{equation}

Because tokenization can split one reasoning step across several adjacent
subtokens, the training signal should use a locally smoothed score rather than
the raw point estimate alone. The default smoothed score is
\begin{equation}
  \bar{S}^{\mathrm{CT}}_{i,t}
  =
  \lambda_0 S^{\mathrm{CT}}_{i,t}
  +
  \lambda_1 S^{\mathrm{CT}}_{i,t-1}
  +
  \lambda_1 S^{\mathrm{CT}}_{i,t+1},
  \qquad
  \lambda_0 + 2\lambda_1 = 1,
\end{equation}
with neighboring terms dropped and the weights renormalized at sequence
boundaries or masked positions. This is not meant to blur across whole
sentences; it is a tokenization-scale regularizer that reduces brittle
single-token spikes, analogous in spirit to smoothing noisy gradient
attributions \cite{smilkov2017smoothgrad}.

The geometric intuition behind this score is that
$\nabla_{h^{(\ell)}_{i,t}}F$ is the normal vector to the local tangent plane of
the training objective with respect to the token-position representation. The
centered activation
\begin{equation}
  \bar{h}^{(\ell)}_{i,t}
  =
  h^{(\ell)}_{i,t}
  -
  \frac{1}{|M_i|}
  \sum_{u \in M_i} h^{(\ell)}_{i,u}
\end{equation}
is a direction from the response's average generated-token state toward the
specific token's representation. Their inner product measures how much the
token-specific representation points in a direction that locally increases or
decreases the scalar objective. Taking the absolute value treats both positive
and negative local influence as importance: a token can matter because it
supports the rewarded trajectory, or because the objective is strongly trying
to move away from its current representation.

This differs from attention in-degree in three important ways. First, a token
must be connected to the objective gradient; receiving attention is not enough.
If $\nabla_{h^{(\ell)}_{i,t}}F \approx 0$, then
$S^{\mathrm{CT}}_{i,t}$ is small even when the token is an attention sink.
Second, centering removes common sequence-level activation structure, so generic
boilerplate magnitude is less likely to dominate. Third, summing over the final
selected layers asks whether the token representation remains useful near the
model's decision surface, rather than whether it was merely read by later
tokens in an earlier attention computation.

For CW-CISPO and AD-CISPO-Bound, the practical interpretation is:
\begin{quote}
  allocate more CISPO learning pressure, and optionally more upper IS clipping
  budget, to tokens whose hidden states lie in directions to which the actual
  policy-gradient objective is locally sensitive.
\end{quote}
This is why causal-tangent saliency is a better match to CISPO extension than
attention centrality. The multiplier is no longer driven by how often a token
is looked at; it is driven by a first-order estimate of how much the training
objective would change if that token-position representation were perturbed.

The proposed next-generation extractor should therefore treat attention
in-degree as a diagnostic baseline and use causal-tangent saliency, plus
explicit sink-token guards, as the first serious replacement when attention
sinks dominate the top-saliency table.

\paragraph{Target policy versus frozen policy.}
Causal-tangent saliency should use activations and gradients from the current
target policy $\pi_\theta$, not from the frozen reference policy
$\pi_{\mathrm{ref}}$. The reason is built into the definition of $F$: it is the
teacher-forced objective whose log-probability terms are produced by the model
being updated. The gradient
\begin{equation}
  \nabla_{h^{(\ell)}_{i,t}} F
\end{equation}
is therefore a sensitivity of the target policy's computation graph. If the
same score were computed on $\pi_{\mathrm{ref}}$, it would estimate the effect
of perturbing hidden states in a model that will not receive the CISPO update.
That would reintroduce the same mismatch that made frozen-reference
attention-saliency fragile: the saliency source could emphasize obsolete
features or sink behavior from a policy that is no longer the object being
optimized.

The frozen policy remains useful as a baseline and diagnostic. It can provide
KL/reference metrics, future-attention in-degree ablations, and drift
measurements. But the primary causal-tangent extractor should run after
rollouts are collected and before the policy update, using the current
training model $\pi_\theta$ on the sampled prompt-response sequences.

\section{Implementation Plan}

The minimum implementation should use the existing Curious training path and
GSM8K reward setup \cite{cobbe2021gsm8k}. The intended experimental model is a
small reasoning-capable open model, such as the repository's Qwen3-1.7B
configuration, so that full ablations are affordable on a single H100.

\begin{center}
\fbox{\begin{minipage}{0.88\linewidth}
\small
\textbf{Next-generation implementation flow:}
$\pi_{\theta_{\mathrm{old}}}$ samples grouped responses
$\rightarrow$ verifier computes rewards and group advantages
$\rightarrow$ target policy is run teacher-forced on sampled responses
$\rightarrow$ final-layer hidden states are captured
$\rightarrow$ the advantage-weighted log-prob objective $F$ is backpropagated
to those hidden states
$\rightarrow$ causal-tangent saliency $S^{\mathrm{CT}}_{i,t}$ is computed and
sink-guarded
$\rightarrow$ saliency and optional entropy are normalized into $w_{i,t}$
$\rightarrow$ optional bound-only ablations also derive $m_{i,t}$
$\rightarrow$ token thresholds
$\epsilon^{IS}_{i,t,\mathrm{high}}=\epsilon^{IS}_{\mathrm{high}}m_{i,t}$
$\rightarrow$ CW-CISPO loss uses detached token weights and detached clipped
IS weights
$\rightarrow$ metrics log reward, exact match, clipping, saliency shape,
credit-weight shape, sink-guard behavior, and runtime overhead.
\end{minipage}}
\end{center}

\paragraph{Implementation contract.}
A mathematically aligned CISPO-extension implementation should satisfy the following:
\begin{enumerate}[leftmargin=*]
  \item CISPO loss must use a detached clipped IS weight multiplying
  $\log\pi_\theta$, not a PPO/GRPO surrogate minimum.
  \item Adaptive CISPO variants must be enabled only with the CISPO loss,
  unless explicitly naming and evaluating an adaptive-GRPO ablation.
  \item MiniMax-style CISPO should set KL weight to zero and avoid an active
  lower IS-weight bound.
  \item Token-adaptive clipping should change only the upper IS clipping radius
  unless a lower-bound ablation is deliberately introduced.
  \item Token-credit weighting should use detached nonnegative weights
  $w_{i,t}$ projected to mean $1$ per response; raw saliency values should not
  be multiplied directly into the loss.
  \item CW-CISPO should first keep scalar CISPO clipping fixed. Bound
  redistribution should be introduced as a separate AD-CISPO-Bound or hybrid
  ablation.
  \item Saliency tensors should align with the generated token whose IS weight
  is being clipped. In next-token prediction implementations, generated-token
  saliency must align with next-token log-probability terms, not the preceding
  hidden-state position.
  \item Causal-tangent saliency should be computed from the target policy on the
  sampled completions, with the scalar objective using the same generated-token
  log-probability terms and group advantages that drive the CISPO update.
  Frozen-policy activations should be reserved for explicit ablations, not for
  the primary extractor.
  \item The causal-tangent backward pass should differentiate only the saliency
  objective used to score token positions; the final CISPO update should still
  use the normal training loss and should not accidentally reuse stale gradients
  from saliency extraction.
  \item Sink guards should be explicit and logged. Separator punctuation,
  whitespace, special tokens, and prompt tokens should either be removed from
  weight and multiplier assignment or capped at $1$, while numeric tokens and
  mathematical operators should remain eligible unless ablated.
  \item Claims about compute and memory must distinguish active-policy saliency
  overhead from frozen-reference overhead.
\end{enumerate}

\paragraph{Causal-tangent extractor steps.}
The recommended first implementation is:
\begin{enumerate}[leftmargin=*]
  \item Add a saliency method flag such as
  \texttt{ad\_cispo\_saliency\_method=causal\_tangent}.
  \item During trajectory processing, run the target policy teacher-forced over
  the sampled prompt-response sequences, using the same logits-minibatch
  boundary used for log-probability computation.
  This should use the current training model before the update step, not the
  frozen reference model.
  \item Register hooks on the last four decoder layers and retain gradients for
  their hidden-state tensors at generated-token positions.
  \item Form $F$ from generated-token log-probabilities weighted by
  $|\hat{A}_i|$, with per-response length normalization so long completions do
  not dominate saliency solely by length.
  \item Backpropagate $F$, compute $S^{\mathrm{CT}}_{i,t}$ from
  gradient--centered-activation inner products, then detach and clear temporary
  gradients before the policy update.
  \item Apply local smoothing, sink guards, and log the fraction of tokens and saliency mass
  removed by those guards.
  \item Add mean-preserving token-credit-weight projection. Reuse the existing
  multiplier projection for bound-only and hybrid ablations, but keep
  $w_{i,t}$ and $m_{i,t}$ separately logged.
\end{enumerate}

\section{Alternative CISPO Extensions}

CW-CISPO is the lead proposal, but the literature review suggests several
other CISPO-native directions that solve problems CISPO has not yet addressed:

\paragraph{1. Entropy-Capacity CISPO.}
CISPO clips IS weights but does not decide which tokens can carry useful
hindsight credit. Entropy-Capacity CISPO would keep the scalar CISPO objective
and weight tokens by a bounded function of token entropy. This is cheaper than
causal-tangent saliency and directly tests the hypothesis that reasoning gains
come from uncertain decision points. Its weakness is that entropy measures
capacity, not whether the token is actually relevant to the rewarded
trajectory.

\paragraph{2. Causal-Token CW-CISPO.}
This is the main proposal: combine causal-tangent saliency with optional entropy
capacity. It solves a more specific problem than entropy-only weighting: not
only which tokens \emph{can} change, but which token-position representations
the current policy-gradient objective is locally sensitive to. Its cost is the
extra teacher-forced backward pass.

\paragraph{3. Selective-CISPO.}
Instead of weighting all tokens, optimize only the top-$p$ percent of tokens by
saliency, entropy, or their product, while still using CISPO's clipped IS
coefficient. This is analogous in spirit to selective preference optimization
\cite{yang2025sepo}, but applied online to RLVR rollouts rather than offline
preference pairs. It could reduce memory and compute, but is riskier because
hard selection can drop useful connective tokens.

\paragraph{4. Variance-Adaptive CISPO.}
CISPO clips IS weights to control variance, but the clipping radius is fixed.
A variance-adaptive extension would tune $\epsilon^{IS}_{\mathrm{high}}$ from
observed ratio dispersion, clipped-token fraction, and advantage variance,
rather than from saliency. This is less novel conceptually, but it may be a
strong practical baseline because recent AD-CISPO runs show that adaptive
saliency bounds rarely bind.

\paragraph{5. Step-Chunk CISPO.}
Instead of assigning weights at the subtoken level, group neighboring tokens
into reasoning chunks using newline, equation, or sentence boundaries and apply
one CISPO credit weight per chunk. This addresses token-sink artifacts and
tokenization noise. It is less mechanistically fine-grained, but may better
match process-supervision intuitions when the reward depends on reasoning
steps rather than isolated subtokens.

The first experimental priority should be Causal-Token CW-CISPO versus
Entropy-Capacity CISPO and scalar CISPO. If causal saliency does not beat
entropy-only weighting, then the causal-tangent extractor is not yet adding
useful credit information beyond uncertainty capacity.

\section{Evaluation Plan}

\paragraph{Datasets and tasks.}
The evaluation should use a staged math-reasoning suite rather than a single
benchmark:
\begin{enumerate}[leftmargin=*]
  \item \textbf{GSM8K} should remain the first target, because it has a clear
  exact-answer reward and is already supported in Curious
  \cite{cobbe2021gsm8k}. It is the right dataset for fast debugging, seed
  sweeps, and early reward-curve comparisons.
  \item \textbf{SVAMP} should be used as an evaluation-only robustness check for
  arithmetic word-problem variations \cite{patel2021svamp}. This helps test
  whether adaptive CISPO improves reasoning robustness rather than merely fitting
  GSM8K surface patterns.
  \item \textbf{MATH and MATH-500} should be the next difficulty tier.
  MATH provides competition-style problems with step-by-step solutions
  \cite{hendrycks2021math}, while MATH-500 is a smaller representative subset
  commonly used for cheaper repeated evaluation \cite{lightman2023verify}.
  These tasks test whether the token-adaptive clipping signal transfers to
  longer derivations and more diverse mathematical domains.
  \item \textbf{AIME/AMC-style held-out sets} should be reserved for final
  stress testing once answer extraction is stable. If broader olympiad-level
  coverage is needed, add the mathematics split of OlympiadBench as an optional
  evaluation-only benchmark \cite{he2024olympiadbench}.
\end{enumerate}
Training should begin with GSM8K and expand only after scalar CISPO and
adaptive CISPO have matched data loading, answer extraction, rollout count, and
compute budgets. Harder datasets should be used primarily to measure transfer,
not to tune the saliency or token-credit heuristic.

\paragraph{Baselines.}
The minimum comparison set is:
\begin{enumerate}[leftmargin=*]
  \item GRPO/PPO-style surrogate baseline.
  \item DAPO-style asymmetric clipping or clip-high baseline.
  \item MiniMax-style CISPO: no KL, no active lower IS bound, scalar upper
  IS radius.
  \item AD-CISPO-Bound with causal-tangent saliency and sink guards.
  \item CW-CISPO with causal-tangent saliency, scalar CISPO clipping, and
  bounded mean-one token weights.
  \item CW-CISPO with entropy-only weights.
  \item CW-CISPO with causal-tangent saliency multiplied by entropy capacity.
  \item CW-CISPO or AD-CISPO-Bound with exact future-attention in-degree
  saliency and KV-norm proxy saliency.
  \item Shuffled-saliency, shuffled-entropy, and uniform-weight controls.
  \item Hard selective-CISPO top-$p$ token ablation.
  \item Optional adaptive-GRPO ablation, clearly separated from AD-CISPO.
\end{enumerate}

\paragraph{Metrics.}
Report both task performance and optimizer behavior:
\begin{itemize}[leftmargin=*]
  \item exact-match accuracy, verifier reward, and pass@k by dataset and
  difficulty tier;
  \item reward-per-token and reward-per-GPU-hour;
  \item upper-clip saturation rate and token update coverage;
  \item entropy, approximate KL to the reference model when logged, and
  response length;
  \item CW-CISPO weight mean, min, max, saturation fraction, and entropy
  concentration;
  \item AD-CISPO multiplier mean, min, max, saliency mean, and saliency
  concentration for bound-only and hybrid variants;
  \item high-saliency token category frequencies, including comma,
  punctuation, numeric, mathematical-operator, and answer-token fractions;
  \item high-weight token category frequencies, separately from high-saliency
  frequencies when entropy gating is enabled;
  \item sink-guard removed-token fraction and removed-saliency-mass fraction;
  \item saliency-extractor overhead in wall-clock time and peak memory,
  separately for target-policy causal-tangent and frozen-reference baselines.
\end{itemize}

\paragraph{Decision rule.}
CW-CISPO should be considered successful only if it improves exact-match or
reward efficiency over scalar CISPO at comparable compute, or if it provides a
clear stability advantage such as lower collapse rate or smoother reward
improvement. A gain that also appears for shuffled saliency, entropy-only
weights, or uniform weights should be treated as evidence for generic
regularization, uncertainty weighting, or changed effective learning rate, not
for the proposed causal-token mechanism. AD-CISPO-Bound should be retained only
if it beats both scalar CISPO and CW-CISPO, because bound-only adaptation is the
weaker mechanistic lever when policy ratios rarely reach the upper bound.

\section{Milestones}

\begin{enumerate}[leftmargin=*]
  \item \textbf{Week 1:} verify MiniMax-style CISPO against GRPO on GSM8K with
  matched model, batch size, rollout count, and reward function.
  \item \textbf{Week 2:} implement detached mean-one token-credit weights in
  the CISPO loss; support uniform, shuffled, entropy-only, causal-tangent, and
  causal-tangent-times-entropy weight sources.
  \item \textbf{Week 3:} run scalar CISPO, AD-CISPO-Bound, CW-CISPO
  causal-tangent, CW-CISPO entropy-only, and CW-CISPO hybrid saliency-entropy
  ablations with matched seeds.
  \item \textbf{Week 4:} analyze reward curves, clipping dynamics, token-weight
  distribution, saliency concentration, sink-guard behavior, runtime overhead,
  and representative trajectory changes.
\end{enumerate}

\section{Risks, Ablations, and Fallbacks}

\paragraph{Risk: attention saliency is not causal.}
Attention weights may not explain model behavior \cite{jain2019attention}.
This is why the next-generation plan treats future-attention in-degree as a
diagnostic baseline and uses causal-tangent saliency as the primary extractor.
The shuffled saliency ablation remains required to verify that any gain is
coming from the token-importance signal rather than from weight or multiplier
variance.

\paragraph{Risk: attention sinks distort saliency.}
Initial and special tokens can attract attention for non-semantic reasons
\cite{xiao2024streamingllm}. The implementation must mask prompt and sink
tokens, use generated-token-only normalization, and report saliency
concentration diagnostics. For causal-tangent saliency, sink guards should be
kept explicit rather than hidden inside the saliency formula so that the removed
token and removed-mass fractions can be audited.

\paragraph{Risk: causal-tangent overhead dominates.}
Causal-tangent saliency adds an extra objective-specific backward pass. The
fallback is to compute it every $N$ rollout batches, use only the final two
decoder layers, reduce the saliency minibatch size, or fall back to KV-norm or
future-attention in-degree for cheaper diagnostic runs.

\paragraph{Risk: saliency weighting changes the effective learning rate.}
Even mean-one token weights can alter gradient variance and interact with
minibatch composition. The fallback is to keep the response-level mean fixed,
track per-minibatch gradient norms, and compare against uniform random weights
with the same marginal distribution. If the random-weight control matches
CW-CISPO, the gain is not evidence for causal-token credit.

\paragraph{Risk: entropy explains the gain.}
If entropy-only CW-CISPO matches causal-tangent CW-CISPO, the saliency
extractor is unnecessary for the next version. In that case, the project should
pivot to Entropy-Capacity CISPO as the simpler CISPO-native extension, while
keeping causal saliency as a diagnostic.

\paragraph{Risk: budget concentration increases variance.}
Large multipliers may overemphasize a small set of tokens. The fallback is a
bounded weight or multiplier projection with post-clamp renormalization, plus
reporting clip-min, clip-max, weight-min, weight-max, and saturation metrics.

\section{Expected Impact}

If successful, this CISPO-extension program would provide evidence that
reasoning-RL trust regions and token-gradient mass should be token-budgeted
rather than uniformly scalar.
The contribution would be useful even if the exact causal-tangent extractor is
later replaced: the broader idea is to keep CISPO's clipped-IS stability while
adding a falsifiable token-credit layer derived from the current policy's own
rollout computation. A negative result would still be informative because it
would test whether CISPO's scalar gradient allocation is already sufficient,
whether entropy alone explains token credit in RLVR, and whether mechanistic
saliency signals are too weak or too noisy for optimizer control.

\begin{thebibliography}{99}

\bibitem{schulman2017ppo}
John Schulman, Filip Wolski, Prafulla Dhariwal, Alec Radford, and Oleg Klimov.
\newblock Proximal Policy Optimization Algorithms.
\newblock \emph{arXiv:1707.06347}, 2017.
\newblock \url{https://arxiv.org/abs/1707.06347}.

\bibitem{ouyang2022instructgpt}
Long Ouyang et al.
\newblock Training language models to follow instructions with human feedback.
\newblock \emph{arXiv:2203.02155}, 2022.
\newblock \url{https://arxiv.org/abs/2203.02155}.

\bibitem{shao2024deepseekmath}
Zhihong Shao et al.
\newblock DeepSeekMath: Pushing the Limits of Mathematical Reasoning in Open
Language Models.
\newblock \emph{arXiv:2402.03300}, 2024.
\newblock \url{https://arxiv.org/abs/2402.03300}.

\bibitem{deepseek2025r1}
DeepSeek-AI.
\newblock DeepSeek-R1: Incentivizing Reasoning Capability in LLMs via
Reinforcement Learning.
\newblock \emph{arXiv:2501.12948}, 2025.
\newblock \url{https://arxiv.org/abs/2501.12948}.

\bibitem{yu2025dapo}
Qiying Yu et al.
\newblock DAPO: An Open-Source LLM Reinforcement Learning System at Scale.
\newblock \emph{arXiv:2503.14476}, 2025.
\newblock \url{https://arxiv.org/abs/2503.14476}.

\bibitem{minimax2025m1}
MiniMax et al.
\newblock MiniMax-M1: Scaling Test-Time Compute Efficiently with Lightning
Attention.
\newblock \emph{arXiv:2506.13585}, 2025.
\newblock \url{https://arxiv.org/abs/2506.13585}.

\bibitem{vaswani2017attention}
Ashish Vaswani et al.
\newblock Attention Is All You Need.
\newblock \emph{arXiv:1706.03762}, 2017.
\newblock \url{https://arxiv.org/abs/1706.03762}.

\bibitem{jain2019attention}
Sarthak Jain and Byron C. Wallace.
\newblock Attention is not Explanation.
\newblock \emph{arXiv:1902.10186}, 2019.
\newblock \url{https://arxiv.org/abs/1902.10186}.

\bibitem{xiao2024streamingllm}
Guangxuan Xiao et al.
\newblock Efficient Streaming Language Models with Attention Sinks.
\newblock \emph{arXiv:2309.17453}, 2024.
\newblock \url{https://arxiv.org/abs/2309.17453}.

\bibitem{meng2022rome}
Kevin Meng, David Bau, Alex Andonian, and Yonatan Belinkov.
\newblock Locating and Editing Factual Associations in GPT.
\newblock \emph{arXiv:2202.05262}, 2022.
\newblock \url{https://arxiv.org/abs/2202.05262}.

\bibitem{wang2022ioi}
Kevin Wang et al.
\newblock Interpretability in the Wild: A Circuit for Indirect Object
Identification in GPT-2 Small.
\newblock \emph{arXiv:2211.00593}, 2022.
\newblock \url{https://arxiv.org/abs/2211.00593}.

\bibitem{syed2023attribution}
Aaquib Syed et al.
\newblock Attribution Patching Outperforms Automated Circuit Discovery.
\newblock \emph{arXiv:2310.10348}, 2023.
\newblock \url{https://arxiv.org/abs/2310.10348}.

\bibitem{kramar2024atpstar}
Janos Kramar, Tom Lieberum, Rohin Shah, and Neel Nanda.
\newblock AtP*: An efficient and scalable method for localizing LLM behaviour
to components.
\newblock \emph{arXiv:2403.00745}, 2024.
\newblock \url{https://arxiv.org/abs/2403.00745}.

\bibitem{sundararajan2017ig}
Mukund Sundararajan, Ankur Taly, and Qiqi Yan.
\newblock Axiomatic Attribution for Deep Networks.
\newblock \emph{Proceedings of ICML}, 2017.
\newblock \url{https://arxiv.org/abs/1703.01365}.

\bibitem{shrikumar2017deeplift}
Avanti Shrikumar, Peyton Greenside, and Anshul Kundaje.
\newblock Learning Important Features Through Propagating Activation
Differences.
\newblock \emph{Proceedings of ICML}, 2017.
\newblock \url{https://arxiv.org/abs/1704.02685}.

\bibitem{smilkov2017smoothgrad}
Daniel Smilkov, Nikhil Thorat, Been Kim, Fernanda Viegas, and Martin
Wattenberg.
\newblock SmoothGrad: removing noise by adding noise.
\newblock \emph{arXiv:1706.03825}, 2017.
\newblock \url{https://arxiv.org/abs/1706.03825}.

\bibitem{arjona2019rudder}
Jose A. Arjona-Medina, Michael Gillhofer, Michael Widrich, Thomas Unterthiner,
Johannes Brandstetter, and Sepp Hochreiter.
\newblock RUDDER: Return Decomposition for Delayed Rewards.
\newblock \emph{Advances in Neural Information Processing Systems}, 2019.
\newblock \url{https://arxiv.org/abs/1806.07857}.

\bibitem{zeng2024tdpo}
Yongcheng Zeng, Guoqing Liu, Weiyu Ma, Ning Yang, Haifeng Zhang, and Jun Wang.
\newblock Token-level Direct Preference Optimization.
\newblock \emph{Proceedings of ICML}, 2024.
\newblock \url{https://arxiv.org/abs/2404.11999}.

\bibitem{liu2025tisdpo}
Aiwei Liu et al.
\newblock TIS-DPO: Token-level Importance Sampling for Direct Preference
Optimization With Estimated Weights.
\newblock \emph{Proceedings of ICLR}, 2025.
\newblock \url{https://arxiv.org/abs/2410.04350}.

\bibitem{yang2025sepo}
Kailai Yang, Zhiwei Liu, Qianqian Xie, Jimin Huang, Erxue Min, and Sophia
Ananiadou.
\newblock Selective Preference Optimization via Token-Level Reward Function
Estimation.
\newblock \emph{Proceedings of EMNLP}, 2025.
\newblock \url{https://arxiv.org/abs/2408.13518}.

\bibitem{zhou2025treg}
Wenxuan Zhou, Shujian Zhang, Lingxiao Zhao, and Tao Meng.
\newblock T-REG: Preference Optimization with Token-Level Reward
Regularization.
\newblock \emph{Proceedings of ACL}, 2025.
\newblock \url{https://aclanthology.org/2025.acl-long.1353/}.

\bibitem{yang2025tidpo}
Ning Yang, Hai Lin, Yibo Liu, Baoliang Tian, Guoqing Liu, and Haijun Zhang.
\newblock Token-Importance Guided Direct Preference Optimization.
\newblock \emph{arXiv:2505.19653}, 2025.
\newblock \url{https://arxiv.org/abs/2505.19653}.

\bibitem{he2026hindsightcredit}
Yuhang He et al.
\newblock Where Hindsight Credit Can Reside: A Signed-Capacity View of Token
Updates in RLVR.
\newblock \emph{arXiv:2604.11056}, 2026.
\newblock \url{https://arxiv.org/abs/2604.11056}.

\bibitem{cobbe2021gsm8k}
Karl Cobbe et al.
\newblock Training Verifiers to Solve Math Word Problems.
\newblock \emph{arXiv:2110.14168}, 2021.
\newblock \url{https://arxiv.org/abs/2110.14168}.

\bibitem{patel2021svamp}
Arkil Patel, Satwik Bhattamishra, and Navin Goyal.
\newblock Are NLP Models really able to Solve Simple Math Word Problems?
\newblock \emph{arXiv:2103.07191}, 2021.
\newblock \url{https://arxiv.org/abs/2103.07191}.

\bibitem{hendrycks2021math}
Dan Hendrycks et al.
\newblock Measuring Mathematical Problem Solving With the MATH Dataset.
\newblock \emph{arXiv:2103.03874}, 2021.
\newblock \url{https://arxiv.org/abs/2103.03874}.

\bibitem{lightman2023verify}
Hunter Lightman et al.
\newblock Let's Verify Step by Step.
\newblock \emph{arXiv:2305.20050}, 2023.
\newblock \url{https://arxiv.org/abs/2305.20050}.

\bibitem{he2024olympiadbench}
Chaoqun He et al.
\newblock OlympiadBench: A Challenging Benchmark for Promoting AGI with
Olympiad-Level Bilingual Multimodal Scientific Problems.
\newblock \emph{arXiv:2402.14008}, 2024.
\newblock \url{https://arxiv.org/abs/2402.14008}.

\end{thebibliography}

\end{document}
